---
title: "Medical Analysis"
author: "Nahom Diress"
date: "`r Sys.Date()`"
output: html_document
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
insurance <- read.csv("insurance.csv")
```

## How the charges are distrubited
Most charges are less than $ 20,000, but there are still significant number of charges that range from 20k to 50k.

```{r,echo=FALSE, message=FALSE, warning = FALSE}
library(ggplot2)
ggplot(insurance, aes(x = charges)) +
  geom_density(fill = "white", color = "black") +
  labs(title = "Distribution of Charges", x = "Charges") 
```

## Charges by age
Generally speaking, as age increases so do medical charges. Here are the data points

```{r,echo=FALSE, message=FALSE, warning = FALSE}
charges_by_age <- insurance %>%
  group_by(age) %>%
  summarize(mean_charges = mean(charges))
ggplot(data = charges_by_age)+
  geom_line(aes(x = age , y = mean_charges))

```

## Average charges by BMI 
Generally speaking, as BMI increases so do medical charges

```{r,echo=FALSE, message=FALSE, warning = FALSE}
insurance <- read.csv("insurance.csv")
library(ggplot2)
library(dplyr)

# Assuming you have loaded your insurance data into a variable named 'insurance'

insurance <- insurance %>%
  mutate(bmi = as.integer(bmi)) %>%
  mutate(bmi_category = cut(bmi,
                            breaks = c(-Inf, 18.5, 25, 30, Inf),
                            labels = c("low", "middle", "high", "very high")))

bmi_charges <- insurance %>%
  group_by(bmi_category) %>%
  summarize(mean_charges = mean(charges))

ggplot(data = bmi_charges, aes(x = bmi_category, y = mean_charges)) +
  geom_bar(stat = "identity") +
  labs(title = "Mean Charges by BMI Category",
       x = "BMI Category",
       y = "Mean Charges") 
```
## Average charges by smoker status
The average charges for smokers are much higher than the average charges for non-smokers
```{r, echo= FALSE, message= FALSE, warning= FALSE}
average_charges_by_smoker <- insurance %>%
  group_by(smoker) %>%
  summarize(avg_charges = mean(charges))

ggplot(data = average_charges_by_smoker, aes(x = smoker, y = avg_charges)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Average Charges by Smoker Status",
       x = "Smoker Status",
       y = "Average Charges")
    ```
    
## Average charges by region 
The average charges were fairly similar across all regions. The average for the southeast was slightly higher.
```{r, echo= FALSE, message= FALSE, warning= FALSE}

average_charges_by_region <- insurance %>%
  group_by(region) %>%
  summarize(avg_charges = mean(charges))

ggplot(data = average_charges_by_region, aes(x = region, y = avg_charges, fill = region)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Average Charges by Region",
       x = "Region",
       y = "Average Charges")

```

# Summary

Here are four insights from this analysis 

- Charges increase with age
 

- Average charges are higher for people with BMI in the "very high" range
  

- Average charges are significantly higher for people who smoke
  
